\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format


% \usepackage{draftwatermark}
% \SetWatermarkText{Draft}
% \SetWatermarkScale{5}
% \SetWatermarkLightness {0.9} 
% \SetWatermarkColor[rgb]{0.7,0,0}


\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps� with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflat						\label{thm:integral_domain}

								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{mathrsfs}
\usepackage[hyphens,spaces,obeyspaces]{url}
\usepackage{url}
\usepackage{hyperref}
\usepackage{subcaption}
\usepackage{authblk}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage[export]{adjustbox}
\usepackage{fixltx2e}
\usepackage{hyperref}
\usepackage{alltt}
\usepackage{color}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{float}
\usepackage{bigints}
\usepackage{braket}
\usepackage{siunitx}


\usepackage[hyphenbreaks]{breakurl}

\newtheorem{thm}{Theorem}[section]
% \newtheorem{defn}[thm]{Definition}
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{lemma}{Lemma}[section]




\newcommand{\veq}{\mathrel{\rotatebox{90}{$=$}}}
\DeclareMathOperator{\bda}{\Big \downarrow}


\DeclareMathOperator{\E}{\mathbb{E}}
\newcommand{\argmax}{\operatornamewithlimits{argmax}}
\newcommand{\argmin}{\operatornamewithlimits{argmin}}

\title{A Few Notes on the Division Algorithm (WIP)}
\author{David Meyer \\ dmm@\{1-4-5.net,uoregon.edu\}}

\date{Last update: \today}							% Activate to display a given date or no date



\begin{document}
\maketitle

\section{Introduction}

TBD


\section{Fields and the Division Algorithm}

\begin{definition}
Let $f$ and $g$ be polynomials in $F[x$]. Then we say that $f$ divides $g$, or $g$ is divisible by $f$ if there is a polynomial $h$ with $g = fh$.
\end{definition}

\noindent
For integers, the greatest common divisor of two integers $a$ and $b$ is the largest integer dividing both
$a$ and $b$. This definition doesn't  quite work for polynomials. In particular, while we cannot talk
about "largest" polynomial in the same manner as we do for integers, we can talk about the degree
of a polynomial. 

\bigskip
\noindent
Recall that the degree of a nonzero polynomial $f$ is the largest integer $m$
for which the coefficient $a_m$ of $x^m$ is nonzero. For example, if $f(x) = a_nx^n + \hdots + a_1x+a_0$ with $a_n \neq 0$, 
then the degree of $f(x)$, written $\deg f$, is $n$. The degree function allows
us to measure size of polynomials. 

\bigskip
\noindent
There is one extra complication with this definition of $\deg f$. Consider for example that any polynomial of the form 
$ax^2$ with $a \neq 0$ divides $x^2$ and $x^3$. Thus, there isn't a unique polynomial of
highest degree that divides a pair of polynomials. To get around this uniqueness problem the convention is to consider 
the \emph{monic} polynomials, that is, those polynomials whose leading coefficient is $1$. For example, $x^2$ is the 
monic polynomial of degree $2$ that divides both $x^2$ and $x^3$, while $5x^2$ is not monic. 

\bigskip
\noindent
Finally, I'll use $ab$ to represent $a \cdot b$ where the meaning is unambiguous. 


\begin{definition}
Let $f$ and $g$ be polynomials over $F$, where $f$ and $g$ are not both zero. Then a greatest common divisor of $f$ and $g$, $\gcd(f,g)$, 
is a monic polynomial of largest degree that divides both $f$ and $g$.
\label{def:gcd}
\end{definition}


\noindent
Unfortunately, there is a problem with Definition \ref{def:gcd} which has to do with uniqueness. In particular, could there be more than one greatest common 
divisor of a pair of polynomials?  As we will se below, the answer turns out to be no.

\bigskip
\noindent
Note that the main reason for assuming that the coefficients of our polynomials lie in a field is to ensure that the division algorithm is valid. Why?
Because division is well behaved in integral domains (where there are no zero divisors), and because every field is an integral domain and every finite 
integral domain is a field. But why is that?


\begin{thm}
Every field $F$ is an integral domain.
\label{thm:integral_domain}
\end{thm}

\noindent
\textbf{Proof:} Recall that if $F$ is a field then each non-zero $r \in F$ has an inverse $r^{-1}$. 
So suppose $r,s \in F$ and $r \neq 0$ such that $r \cdot s = 0$. Then the claim is that $s = 0$.
Why? Consider


\begin{equation}
\begin{array}{rcll} 
r \cdot s 
&=& 0                                                                                &\quad  \mathrel{\#} \text{assumption with $r \neq 0$}         \\
&\Rightarrow& r^{-1} \cdot (r \cdot s) = r^{-1} \cdot 0        &\quad  \mathrel{\#} \text{multiply both sides by $r^{-1}$}    \\
&\Rightarrow&  r^{-1} \cdot (r \cdot s) =  0                        &\quad  \mathrel{\#} x \cdot 0 = 0                                          \\
&\Rightarrow&  (r^{-1} \cdot r) \cdot s =  0                        &\quad  \mathrel{\#} \text{multiplication is associative}         \\
&\Rightarrow&  1 \cdot s = 0                                             &\quad  \mathrel{\#} r^{-1} \cdot r = 1                                   \\
&\Rightarrow&  s = 0                                                         &\quad  \mathrel{\#} 1 \cdot x = x
\end{array}
\end{equation}

\bigskip
\noindent
So $r$ is not a zero divisor. But every non-zero element $r$ of the field $F$ has an inverse ($r$ is a "unit")
so $F$ has no zero divisors and is by definition an integral 
domain. $\square$

\bigskip
\begin{thm}
Every finite integral domain is a field.
\label{thm:finite_id_is_a_field}
\end{thm}

\noindent
\textbf{Proof:}  The proof is based on the fact that since $R$ is an integral domain it has a cancellation law (or equivalently, $R$ has no zero divisors). 
Having a cancellation law means that

\begin{equation}
ab = ac \implies b = c
\label{eqn:cancellation_law}
\end{equation}


\bigskip
\noindent
To see why any finite integral domain $R$ is a field, consider $R = \{r, r^2, r^3, \hdots, r^n\}$  where 
$r^k \neq 0$ for $1 \leq k \leq n$.  Since $R$ is finite we will have $r^k = r^l$ for some $k$ and $l$ such that $k > l$. Then

\bigskip
\begin{equation*}
\begin{array}{rlll} 
r^k 
&=& r^l                                                                                        &    \qquad \mathrel{\#} \text{$R$ is a finite integral domain}        \\
&\Rightarrow& r \cdot r^{k-1} = r \cdot r^{l -1}                             &    \qquad \mathrel{\#} \text{factor out $r$}                                  \\
&\Rightarrow& r^{k-1} = r^{l -1}                                                   &    \qquad\mathrel{\#} \text{use cancellation law (cancel $r$, Equation \ref{eqn:cancellation_law}})    \\
&\Rightarrow& r \cdot r^{k-2} = r \cdot r^{l -2}                             &    \qquad\mathrel{\#} \text{factor out $r$}                                   \\
&\Rightarrow& r^{k-2} = r^{l -2}                                                   &    \qquad\mathrel{\#} \text{use cancellation law (cancel $r$, Equation \ref{eqn:cancellation_law})}  \\
&\vdots                                                                                        && \qquad\mathrel{\#} \text{iterate $l - 1$ times}                          \\    
&\Rightarrow& r^{k - l + 1} = r^1                                                 &     \qquad\mathrel{\#} \text{...}                                                    \\       
&\Rightarrow& r \cdot r^{k - l} = r \cdot r^0                                 &     \qquad\mathrel{\#} \text{factor out $r$}                                   \\    
&\Rightarrow& r^{k-l} = r^{0}                                                       &    \qquad\mathrel{\#} \text{use cancellation law (cancel $r$, Equation \ref{eqn:cancellation_law})}   \\   
&\Rightarrow& r^{k-l} = 1                                                            &     \qquad\mathrel{\#} r^0 = 1
\end{array}
\end{equation*}

\bigskip
\noindent
So $ r^{k-l} = 1 $. If $k-l = 1$ then $r$ a unit since $r^{k-l} = r^{1} = 1$ so $r^{-1}$ is $\frac{1}{r}$. Otherwise
$k-l > 1$ and $r^{k-l} = 1 \Rightarrow r^{k-l-1} = \frac{1}{r}$.   So $r^{-1} = r^{k-l -1}$ and 
every $r \neq 0 \in R$ has an inverse. Thus every non-zero  $r \in R$ is a unit and so $R$ is a field. $\square$

\bigskip
\noindent
So every field is an integral domain and every finite integral domain is a field and so division behaves reasonably in these cases. 


\bigskip
\noindent
A few conventions: set $\deg 0 = -\infty$ and $-\infty + -\infty = - \infty$ and $-\infty +n = -\infty$ for $n \in \mathbb{Z}$.

\begin{lemma}
Let $F$ be a field and let $f$ and $g$ be polynomials over $F$. Then $\deg fg = \deg f  + \deg g$.
\label{lemma:degree}
\end{lemma}

\noindent
\textbf{Proof:} If either $f = 0$ or $g = 0$, then the equality $\deg fg = \deg f + \deg g$ is true by our convention above. So, suppose that 
$f \neq 0$ and $g \neq 0$. Write $f = a_n x^n + \hdots + a_0$ and $g = b_m x^m + \hdots + b_0$ with $a_n \neq 0$ and $b_m \neq 0$.  
Therefore, $\deg f = n$ and $\deg g = m$. The definition of polynomial multiplication yields

\begin{equation*}
fg=(a_nb_m)x^{n+m} +(a_nb_{m-1} + a_{n-1} b_m)  x^{n+m-1} + \hdots +a_0b_0
\end{equation*}

\bigskip
\noindent
Since every field is an integral domain and since the coefficients come from a field we know that there are no zero divisors and so we can conclude
that $a_nb_m \neq 0$, and so $\deg fg = n + m = \deg f+ \deg g$, as desired. $\square$

% \newpage


\begin{definition}
The characteristic of a field $F$, $\text{char}(F)$, is the smallest positive integer $n$ such that

\begin{equation*}
n \cdot 1 = \underbrace{1 + 1 + \cdots + 1}_{n \text{ times}} = 0
\end{equation*}
\end{definition}



\bigskip
\noindent
The characteristic has all kinds of interesting properties. For example

\begin{lemma}
Let $F$ be a field. Then
\begin{enumerate}
\item If $\text{char}(F) > 0$  then $\text{char} (F)$ is prime
\label{char:1}
\item If $F$ is a finite field then $\text{char} (F)  >  0$
\label{char:2}
\item If $F$ is a finite field then $\text{char} (F)$ is prime
\end{enumerate}
\end{lemma}

\bigskip
\noindent
\textbf{Proof:} 

\begin{enumerate}
\item Assume that $n$ is not prime so that there exists a nontrivial factorization of $\text{char}(F) = n = p \cdot  q$. Then

\begin{equation*}
\begin{array}{rcll}
0
&=&  n \cdot 1                                                           &\qquad \mathrel{\#} \text{definition of $\text{char}(F)$} \\
&=&  (p \cdot q) \cdot 1                                             &\qquad \mathrel{\#} n = p \cdot q \\
&=&  p \cdot (q \cdot 1)                                             &\qquad \mathrel{\#} \text{multiplication is associative} \\
&=& (p \cdot 1) \cdot (q \cdot 1)                                &\qquad \mathrel{\#} p = p \cdot 1
\end{array}
\end{equation*}

So $(p \cdot 1) \cdot (q \cdot 1) = 0$. But we know that $F$ is a field and fields have no zero divisors so either $(p \cdot 1) = 0$ or 
$(q \cdot 1) = 0$. But this contradicts the assumption that  $\text{char}(F)$ was the smallest positive integer $n$ such that $n \cdot 1 = 0$ 
since $p, q \neq 0$ and $n = p \cdot q$. So $n = \text{char}(F)$ must have been prime.
\label{enumerate:1}

Another way to think about this proof is to notice that if $F$ is a field of characteristic $n$ then the elements $\{0,1,2,\hdots,,n-1\}$ of $F$ 
obey the rules for addition and  multiplication modulo $n$. Therefore there is a copy of $\mathbb{Z}_n$ inside of $F$. Since $\mathbb{Z}_n$ 
has zero divisors when $n$ is not prime (and a field is an integral domain and so has no zero divisors, see Theorem \ref{thm:integral_domain}), 
it follows that the characteristic of a field must be prime. 

Thus every finite field $F$ must have characteristic $p$ for some prime $p$ and the elements $\{0,1,2,\hdots,p-1\}$ form a copy of 
$\mathbb{Z}_p$ inside of $F$. This copy of $\mathbb{Z}_p$ is known as the \emph{prime subfield of F}.



\item Since $F$ is finite not every element can be unique. In particular,  consider a finite field $F = \{f, f^2, f^3, \hdots, f^n\}$  where 
$f^k \neq 0$ for $1 \leq k \leq n$.  Since $F$ is finite we will have $(f^k \cdot 1) = (f^l \cdot 1)$ for some $k$ and $l$ where $k > l$. 
Then $(f^k -  f^l ) \cdot 1  = 0$ and so $\text{char}(F) > 0$.
\label{enumerate:2}

\item Immediate by \ref{enumerate:1}. and \ref{enumerate:2}.
\end{enumerate}
\noindent
$\square$



\begin{proposition} \textbf{(Division Algorithm)}
Let $F$ be a field and let $f$ and $g$ be polynomials over $F$ with $f$ nonzero. Then there are unique polynomials $q$ and $r$ with $g = qf + r$
with $\deg r  < \deg f$.
\end{proposition}

\noindent
\textbf{Proof:}  Let $\mathcal{S} = \{t \in F[x] \mid t = g - qf \text{ for 
some } q \in F[x] \}$. Now, we know $\mathcal{S} \neq \emptyset$
since $g \in \mathcal{S}$ (see Equation \ref{eqn:f} below). We also know by the well ordering property
of the integers \cite{well_ordering_principle} that there is a polynomial $r$ of least degree in $\mathcal{S}$. Then
by definition there is a $q \in F[x]$ with $r = g - qf$ and so $g = qf +r$. 

\bigskip
\noindent
We need to show that $\deg r < \deg f$. So suppose that $\deg r \geq \deg f$, say with  $n = \deg f$ and $m = \deg r$. 
If $f = a_nx^n + \hdots +a_0$ and
$r = r_mx^m + \hdots +r_0$ with $a_n \neq 0$ and $r_m \neq 0$, then by the method of
long division of polynomials, we see that $r = (r_m a_n^{-1})x^{m - n} f + r^\prime$ with  $\deg r^\prime < m = \deg r$. But then 


\begin{equation*}
\begin{array}{rcll}
g
&=& qf + r                                                                  &\quad  \mathrel{\#} g \in \mathcal{S} \text{ and well ordering principle} \\
&=& qf + (r_m a_n^{-1}) x^{m - n}f + r^\prime           &\quad  \mathrel{\#}  r = (r_m a_n^{-1})x^{m - n} f + r^\prime \\
&=& (q + r_m a_n^{-1} x^{m - n}) f + r^\prime           &\quad  \mathrel{\#} \text{factor out $f$} \\
&\implies& r^\prime \in \mathcal{S}                           &\quad  \mathrel{\#} r^\prime = g - (q + r_m a_n^{-1} x^{m - n}) f \in \mathcal{S}
\end{array}
\end{equation*}

\bigskip
\noindent
Since $\deg r^\prime < \deg r$,   $r^\prime \in \mathcal{S} $ contradicts our choice of $r$ (the assumption that $r$ was the polynomial of least degree
in $\mathcal{S}$).  Therefore $\deg r \geq \deg f$ is false so $\deg r  < \deg f$ is true and this proves existence of $q$ and $r$.

\bigskip
\noindent
We still have to show the uniqueness of $q$ and $r$. To do this, suppose that $g = qf +r$ and $g = q^{\prime} f + r^\prime$ for some polynomials 
$q,q^\prime, r,r^\prime \in F[x]$, and with both $\deg r < \deg f$ and $\deg r^\prime <  \deg f$. Then $qf +r = q^{\prime} f +r^{\prime}$, so 
$(q -  q^{\prime})f = r^\prime - r$. Taking degrees and using Lemma \ref{lemma:degree}, we have 

\begin{equation}
\deg (q - q^{\prime}) + \deg f  = \deg (r^\prime - r)
\label{eqn:qminusqprime}
\end{equation}

\bigskip
\noindent
Since $\deg r <  \deg f$ and $\deg r^\prime <  \deg f$, we have $\deg (r^\prime - r) < 0$. However, if $\deg(q^\prime - q) \geq 0$,  
we get a contradiction of Equation \ref{eqn:qminusqprime}. In particular, by our convention the only way for this to happen is for 
$\deg(q^\prime - q) = \deg(r^\prime - r) = - \infty$, and the only way for that to happen is if $q^\prime - q = 0 = r^\prime - r$. So $q^\prime = q$ and  
$r^\prime =r$, which shows uniqueness. $\square$

\bigskip
\begin{proposition}
Let $F$ be a field and let $f$ and $g$ be polynomials over $F$, with not both $f$ and $g$ zero. Then $\gcd(f, g)$ exists and is unique. 
Furthermore, there are polynomials $h$ and $k$ with $\gcd(f, g) = hf + kg$.
\end{proposition}

\noindent
\textbf{Proof:}  Here we have to show that $\gcd(f, g)$ exists and that it is unique. To show that $\gcd(f, g)$ exists, 
let $\mathcal{S} = \{hf + kg : h, k \in F [x]\}$. Then $\mathcal{S}$ contains nonzero polynomials $f$ and $g$ (at least), since
for some $f^\prime \in \mathcal{S}$

\begin{equation}
\begin{array}{rcll}
f^\prime
&=& hf +  kg                                            &\quad  \mathrel{\#} \text{definition of elements of } \mathcal{S}\\
&=& 1 \cdot f + 0 \cdot g                         &\quad  \mathrel{\#} \text{let $h = 1$ and $k = 0$}  \\
&=& f                                                       &\quad  \mathrel{\#} \text{so $f \in \mathcal{S}$}                    \\
\end{array}
\label{eqn:f}
\end{equation}

\bigskip
\noindent
Likewise for $g$ (let $h = 0$ and $k = 1$). So $f,g \in \mathcal{S}$. But this means that there is a nonzero polynomial $d \in \mathcal{S}$ of 
smallest degree by the well ordering principle \cite{well_ordering_principle}, so we can write $d = hf + kg$ for some $h, k \in F [x]$. By dividing by the leading coefficient of 
$d$, we may assume that $d$ is monic without changing the  condition that $d$ is the polynomial of least degree in $\mathcal{S}$.

\bigskip
\noindent
Now the claim is that $d = \gcd(f,g)$. To show that $d$ is a common divisor of $f$ and $g$, first consider $f$. By the division algorithm, we can
write $f$ as $f = qd + r$ for some polynomials $q$ and $r$ with $\deg r < \deg d$. Then

\begin{equation}
\begin{array}{rcll}
f
&=& qd +  r                                            &\mathrel{\#}  f \in \mathcal{S} \text{ by Equation \ref{eqn:f}}, \mathcal{S} \text{ has a division algorithm} \\
&\implies & r = f  - qd                             &\mathrel{\#} \text{solve for $r$}\\
&\implies& r = f - q(hf + kg)                    &\mathrel{\#}  d = hf + kg \\
&\implies& r = f - qhf  - qkg                     &\mathrel{\#} \text{arithmetic}\\
&\implies& r = (1 - qh)f  + (-qk)g             &\mathrel{\#} F[x]  \text{ is closed under */+ so } r  \text{ is of the form } hf + kg \\
&\implies&  r \in \mathcal{S}                   &\mathrel{\#} r \text{ is of the form } hf + kg \text{ so } r \in \mathcal{S} 
\end{array}
\label{eqn:division}
\end{equation}


\bigskip
\noindent
So $r \in \mathcal{S}$. Now, if $r \neq 0$ we have contradiction since we assumed that  $\deg(r) < \deg(d)$ and that $d$ was the polynomial of lowest degree
in $\mathcal{S}$. Thus $r$ must equal $0$, which shows that $f = qd$, and so $d$ divides $f$. Similarly, $d$ divides $g$ and so $d$ is a common divisor of $f$ and $g$.

\bigskip
\noindent
If  $e$ is any other common divisor of $f$ and $g$, then e divides any combination of $f$ and $g$; in particular, $e$ divides $d= hf + kg $. 
This forces $\deg(e) \leq \deg(d)$ by Lemma \ref{lemma:degree}. Thus, $d$ is the monic polynomial of largest degree that divides $f$ and $g$, so $d$ is a greatest 
common divisor of $f$ and $g$.  This shows that $\gcd(f,g)$ exists.

\bigskip
\noindent
To show that $\gcd(f,g)$ is unique, suppose that $d$ and $d^\prime$ are both monic common divisors of $f$ and 
$g$ of largest degree.  By Equation \ref{eqn:division}, we can write both $d$ and $d^\prime$ as combinations of $f$ and $g$. In addition, the argument above shows that 
$d$ divides $d^\prime$ and vice-versa. If $d^\prime = ad$ and $d = bd^\prime$, then $d = bd^\prime = bad$. Taking degrees shows that $\deg(ba) = 0$, which means 
that $a$ and $b$ are both constants. However since $d$ and $d^\prime$ are monic, for $d^\prime = ad$ to be monic it must be that $a = 1$. Thus, $d^\prime = ad = d$. 
This shows that the greatest common divisor is unique, completing the proof.


\bigskip
\section{Acknowledgements}

\newpage
\bibliographystyle{plain}
\bibliography{/Users/dmm/papers/bib/qc}
\end{document} 
